KIN: A method to infer relatedness from low-coverage ancient DNA

### 2022-09-07 paper_overleaf/KIN_a_method_to_estimate_relatedness_from_low_coverage_data_supplements.tex
Updates from Overleaf

>>>>

\begin{figure}[h]
    \centering
    \includegraphics[width=18cm]{supplementary_info/plots/contam_missspecification.png}
    \caption{False positive and true positive rates as a function of incorrect contamination estimates provided to KINgaroo. Dashed vertical line represents true contamination (2.5\%) in simulations, and dotted horizontal line in top panel represents $5\%$ false positive rate.}
    \label{figS10:cnt}
\end{figure}

### 2022-09-07 paper_overleaf/main.tex

Contamination can affect the accuracy of relatedness inference for two reasons. For one, if a substantial fraction of samples is contaminated, then the estimation of $p_0$ becomes inaccurate, because the majority of pairwise differences will include at least one contaminated sample. The second issue is that contaminated samples will look less similar to other individuals, and thus cause a bias towards inferring them to be less related to other individuals. The amount of contamination does not need to be large for this to be important, in our simulations we find that even at contamination levels $\leq 3\%$, the performance of READ is substantially reduced. Assuming the contamination rates are correctly inferred, the correction we implemented leads to improved performance compared to naive methods such as READ, although they too could be ameliorated in a similar way \cite{laurits_skov_genetic_nodate}.

>>>>

Contamination can affect the accuracy of relatedness inference for two reasons. For one, if a substantial fraction of samples is contaminated, then the estimation of $p_0$ becomes inaccurate, because the majority of pairwise differences will include at least one contaminated sample. The second issue is that contaminated samples will look less similar to other individuals, and thus cause a bias towards inferring them to be less related to other individuals. The amount of contamination does not need to be large for this to be important, in our simulations we find that even at contamination levels $\leq 3\%$, the performance of READ is substantially reduced. When contamination rates are correctly inferred, the correction we implemented leads to improved performance compared to naive methods such as READ, although they too could be ameliorated in a similar way \cite{laurits_skov_genetic_nodate}. However, in many cases contamination estimates may be uncertain or inaccurate, and we show in fig. S13 that KIN's performance is robust to small deviations (small compared to average pairwise heterozygosity) in contamination estimates.
